Supermassive black holes and their host spheroids \\ II. The red and blue sequence in the $M_{\rm BH} - M_{\rm *,sph}$ diagram

### 2015-11-20 papers/MbhMsph/ApJ_submission/mm.tex
victory2

%\and

%\author{A. W. Graham\altaffilmark{1}}
%\affil{Centre for Astrophysics and Supercomputing, Swinburne University of Technology, Hawthorn, Victoria 3122, Australia.}

%% Notice that each of these authors has alternate affiliations, which
%% are identified by the \altaffilmark after each name.  Specify alternate
%% affiliation information with \altaffiltext, with one command per each
%% affiliation.

%\altaffiltext{1}{}
%\altaffiltext{2}{Society of Fellows, Harvard University.}
%\altaffiltext{3}{present address: Center for Astrophysics,
%    60 Garden Street, Cambridge, MA 02138}
%\altaffiltext{4}{Visiting Programmer, Space Telescope Science Institute}
%\altaffiltext{5}{Patron, Alonso's Bar and Grill}

%% Mark off your abstract in the ``abstract'' environment. In the manuscript
%% style, abstract will output a Received/Accepted line after the
%% title and affiliation information. No date will appear since the author
%% does not have this information. The dates will be filled in by the
%% editorial office after submission.

\begin{abstract}
In our first paper, we performed a detailed (i.e.~bulge, disks, bars, spiral arms, rings, halo, nucleus, etc.) 
decomposition of 66 galaxies, with directly measured black hole masses, 
that had been imaged at $3.6\rm~\mu m$ with \emph{Spitzer}.
Our sample is the largest to date and, for the first time, the decompositions were checked for consistency with the galaxy kinematics. 
We present correlations between the black hole mass, $M_{\rm BH}$, 
and the host spheroid (and galaxy) luminosity, $L_{\rm sph}$ (and $L_{\rm gal}$), 
and also stellar mass, $M_{\rm *,sph}$.
Most previous studies have used galaxy samples that were overwhelmingly dominated by high-mass, early-type galaxies.
Instead, our sample includes 17 spiral galaxies, half of which have $M_{\rm BH} < 10^7\rm~M_\odot$, 
and allows us to better investigate the poorly studied low-mass end of the $M_{\rm BH} - M_{\rm *,sph}$ correlation.
The bulges of early-type (E + S0) galaxies follow $M_{\rm BH} \propto M_{\rm *,sph}^{1.04 \pm 0.10}$  
and define a tight %\emph{early-type sequence} 
\emph{red sequence} with intrinsic scatter $\epsilon_{(Y|X)} = 0.43 \pm 0.06\rm~dex$ 
and a median $M_{\rm BH}/M_{\rm *,sph}$ ratio of $0.68 \pm 0.04\%$, 
i.e.~a $\pm 2\sigma$ range of 0.1--5\%.
At the low-mass end, the bulges of late-type (Sp) galaxies define a much steeper %\emph{late-type sequence}, 
\emph{blue sequence}, 
with $M_{\rm BH} \propto M_{\rm *,sph}^{2-3}$, 
indicating that gas-rich processes feed the black hole more efficiently than the host bulge as they coevolve. 
We additionally report that: i) S\'ersic galaxies follow $M_{\rm BH} \propto M_{\rm *,sph}^{1.48 \pm 0.20}$, a less steep sequence than previously reported; 
ii) bulges with S\'ersic index $n_{\rm sph}<2$, argued by some to be pseudo-bulges, 
are not offset to lower $M_{\rm BH}$ from the correlation defined by the current bulge sample with $n_{\rm sph}>2$; 
and iii) $L_{\rm sph}$ and $L_{\rm gal}$ correlate equally well with $M_{\rm BH}$, in terms of intrinsic scatter, only for early-type galaxies 
(most elliptical galaxies contain disks, i.e.~$L_{\rm sph} \neq L_{\rm gal}$) --
once reasonable numbers of spiral galaxies are included, the correlation with $L_{\rm sph}$ is better than that with $L_{\rm gal}$.

@@

Here we investigate in more detail %substructure in 

@@

\subsection{Black hole mass -- spheroid stellar mass}
Finally, we present the $M_{\rm BH} - M_{\rm *,sph}$ diagram in Figure \ref{fig:mbhmasssph}, 
and its linear regression analysis in Table \ref{tab:lregmass}. 
The bulges of the early-type galaxies follow $M_{\rm BH} \propto M_{\rm *,sph}^{1.04 \pm 0.10}$,
consistent with a dry-merging formation scenario\footnote{\bf In dry mergers, the black hole and the bulge grow at the same pace, 
increasing their mass in lock step.},
and define a tight \emph{early-type sequence} with intrinsic scatter $\epsilon_{(Y|X)} = 0.43 \pm 0.06\rm~dex$. 
\cite{graham2012bent} reported that the $M_{\rm BH}/M_{\rm dyn,sph}$ ratio for core-S\'ersic galaxies was 0.36\% 
($M_{\rm dyn,sph}$ is the spheroid dynamical mass), 
and discussed the many implications of this.  
Using a larger data sample, \cite{grahamscott2013} reported that the $M_{\rm BH}/M_{\rm *,sph}$ ratio was 0.49\% for core-S\'ersic galaxies.   
Here we find a median $M_{\rm BH}/M_{\rm *,sph}$ ratio of $0.50 \pm 0.04\%$ for the 22 core-S\'ersic galaxies 
and $0.68 \pm 0.04\%$ for the 45 early-type galaxies.  
Among other things, this higher value (previously reported to be $0.1 - 0.2\%$ for all galaxy types, e.g.~\citealt{marconihunt2003}), 
boosts estimates of the black hole mass function and mass density based on galaxy/spheroid luminosity functions. \\
The bulges of the spiral galaxies trace a steeper \emph{late-type sequence}, 
whose slope is less well constrained due to the smaller size of the subsample and, more importantly, 
to the smaller range in $M_{\rm *,sph}$ that the subsample spans. 
For the bulges of spiral galaxies, the BCES code returns a log-linear relation with a slope $= 3.00 \pm 1.30$, 
while the modified FITEXY routine finds a shallower (but still consistent within the $1\sigma$ uncertainty) 
slope $= 2.28^{+1.67}_{-1.01}$.
More data would be welcome to better constrain the slope of this \emph{late-type sequence}, 
although we note that direct measurements of black hole masses below $10^6\rm~M_\odot$ are extremely challenging to obtain 
with the current technological resources. 
In this regard, using a sample of $\sim$140 low-redshift ($z \leq 0.35$, with a median redshift $\langle z \rangle = 0.085$) 
bulges hosting Active Galactic Nuclei (AGNs) with virial
black hole masses $10^5 \lesssim M_{\rm BH}/{\rm M_\odot} \lesssim 2 \times 10^6$ \citep{jiang2011a}, 
\cite{grahamscott2015} showed that they roughly follow the quadratic $M_{\rm BH} - M_{\rm *,sph}$ relation defined by their S\'ersic bulges.
The majority of our spiral galaxies host an AGN\footnote{According to the nuclear classification reported on NED 
(NASA Extragalactic Database), among our 17 spiral galaxies, at least 12 host a Seyfert AGN and one hosts a LINER AGN.} and
we anticipate here that the correlation traced by our spiral galaxy bulges 
may track the location of these lower mass AGN in the $M_{\rm BH} - M_{\rm *,sph}$ diagram.
That is, the AGNs appear to be the low-mass continuation of our tentative \emph{late-type sequence} shown in Figure \ref{fig:mbhmasssph} 
and this will be explored with more rigour in a forthcoming paper.

@@

\begin{itemize}
\item The logarithmic $M_{\rm BH} - M_{\rm *,sph}$ relation for the spheroidal components of early-type (elliptical + lenticular) galaxies 
      has a slope of $1.04 \pm 0.10$ and intrinsic scatter $\epsilon_{(Y|X)} = 0.43 \pm 0.06\rm~dex$. 
      We call this tight correlation an \emph{early-type sequence}. 
      The $M_{\rm BH} - M_{\rm *,sph}$ log-relation for the bulges of late-type (spiral) galaxies has a slope of $2-3$,
      which is less well constrained due to
      the smaller size of the subsample and, more importantly, the smaller range in spheroid stellar mass 
      ($3 \times 10^9 \lesssim M_{\rm *,sph}/{\rm M_\odot} \lesssim 3 \times 10^{10}$) that the subsample spans. 
      We refer to this correlation as a \emph{late-type sequence}. 
      In (gas-poor) early-type galaxies, the black hole and the stellar content of the spheroidal component grow at the same pace, 
      following a linear $M_{\rm BH} - M_{\rm *,sph}$ relation. 
      In (gas-rich) spiral galaxies, the black hole grows faster than its host bulge, 
      following a quadratic/cubic $M_{\rm BH} - M_{\rm *,sph}$ relation. 
      Unsurprisingly, in a color-magnitude diagram\footnote{Total $B-V$ colors, 
      corrected for inclination, Galactic extinction and K-correction, 
      were taken from the HyperLEDA online database \citep{hyperleda}.}, 
      our early- and late-type galaxies occupy the two distinct regions of the red sequence and the blue cloud, respectively. 
      For analogy with this, we refer to our \emph{early-type sequence} as a \emph{red sequence} 
      and to our \emph{late-type sequence} as a \emph{blue sequence}. 
\item The median $M_{\rm BH}/M_{\rm *,sph}$ ratio for the early-type galaxies is $0.68 \pm 0.04\%$. 
      This value is dramatically larger than what was previously reported ($0.1 - 0.2\%$ for all galaxy types, e.g.~\citealt{marconihunt2003}), 
      but in close agreement with the value of $0.49\%$ reported by \cite{grahamscott2013} for core-S\'ersic spheroids. 
\item The logarithmic $M_{\rm BH} - M_{\rm *,sph}$ relations for the core-S\'ersic and S\'ersic spheroids 
      have slopes with over-lapping uncertainties ($1.19 \pm 0.23$ and $1.48 \pm 0.20$, respectively).  
      The S\'ersic relation is less steep than, but also has over-lapping uncertainties with, 
      the slope of $2.22 \pm 0.58$ reported by \cite{scott2013} for S\'ersic spheroids.  
      The distinction between core-S\'ersic and S\'ersic spheroids found by \cite{scott2013} is thus less pronounced here. 
\item In the $M_{\rm BH} - L_{\rm sph}$ (or $M_{\rm BH} - M_{\rm *,sph}$) diagram, for early-type galaxies, 
      we did not observe the change in slope required for consistency with the log-linear $M_{\rm BH} - \sigma$ and bent $L_{\rm sph} - \sigma$ correlations.
      This issue of inconsistency remains therefore an open question. 
      It might be that we are still not probing enough low-mass black holes ($M_{\rm BH} < 10^7\rm~M_\odot$) 
      for the subsample of early-type galaxies, 
      or that the transition from  $L_{\rm sph} \propto \sigma^2$ at low luminosities to $L_{\rm sph} \propto \sigma^{(5-6)}$ at high luminosities 
      is less sharp than previously thought.
      We intend to investigate this point in our future work. 
\item Spheroids with S\'ersic index $n_{\rm sph}<2$, argued by some to be pseudo-bulges, 
      are not offset to lower $M_{\rm BH}$ from the $M_{\rm BH} - L_{\rm sph}$ correlation defined by spheroids with $n_{\rm sph}>2$. 
\item The $M_{\rm BH} - L_{\rm gal}$ and $M_{\rm BH} - L_{\rm sph}$ correlations have the same level of intrinsic scatter 
      when considering early-type galaxies only. 
      Once reasonable numbers of spiral galaxies are included, 
      $M_{\rm BH}$ correlates better with $L_{\rm sph}$ than with $L_{\rm gal}$ {\bf (see also \citealt{erwingadotti2012,beifiori2012})}. 
\end{itemize}

@@

\acknowledgments
GS warmly thanks Luca Cortese, Elisabete Lima Da Cunha, Duncan Forbes and Gonzalo Diaz for useful discussion. 
{\bf We thank the anonymous referee for their thorough review and highly appreciate the comments and suggestions, 
which significantly contributed to improving the quality of the publication. } \\
This research was supported by Australian Research Council funding through grants
DP110103509 and FT110100263.
This work is based on observations made with the IRAC instrument \citep{fazio2004IRAC} 
on-board the Spitzer Space Telescope, 
which is operated by the Jet Propulsion Laboratory, 
California Institute of Technology under a contract with NASA.
This research has made use of the GOLDMine database \citep{goldmine} and the NASA/IPAC Extragalactic Database (NED) 
which is operated by the Jet Propulsion Laboratory, California Institute of Technology, 
under contract with the National Aeronautics and Space Administration. 
We acknowledge the usage of the HyperLeda database (\url{http://leda.univ-lyon1.fr}).
The BCES routine \citep{akritasbershady1996} was run via the python module 
written by Rodrigo Nemmen \citep{nemmen2012}, which is available at \url{https://github.com/rsnemmen/BCES}.

>>>>

\begin{abstract}
In our first paper, we performed a detailed (i.e.~bulge, disks, bars, spiral arms, rings, halo, nucleus, etc.) 
decomposition of 66 galaxies, with directly measured black hole masses, $M_{\rm BH}$, 
that had been imaged at $3.6\rm~\mu m$ with \emph{Spitzer}.
Our sample is the largest to date and, for the first time, the decompositions were checked for consistency with the galaxy kinematics. 
We present correlations between $M_{\rm BH}$ 
and the host spheroid (and galaxy) luminosity, $L_{\rm sph}$ (and $L_{\rm gal}$), 
and also stellar mass, $M_{\rm *,sph}$.
While most previous studies have used galaxy samples that were overwhelmingly dominated by high-mass, early-type galaxies,
our sample includes 17 spiral galaxies, half of which have $M_{\rm BH} < 10^7\rm~M_\odot$, 
and allows us to better investigate the poorly studied low-mass end of the $M_{\rm BH} - M_{\rm *,sph}$ correlation.
The bulges of early-type galaxies follow $M_{\rm BH} \propto M_{\rm *,sph}^{1.04 \pm 0.10}$  
and define a tight 
\emph{red sequence} with intrinsic scatter $\epsilon_{(M_{\rm BH}|M_{\rm *,sph})} = 0.43 \pm 0.06\rm~dex$ 
and a median $M_{\rm BH}/M_{\rm *,sph}$ ratio of $0.68 \pm 0.04\%$, 
i.e.~a $\pm 2\sigma$ range of 0.1--5\%.
At the low-mass end, the bulges of late-type galaxies define a much steeper 
\emph{blue sequence}, 
with $M_{\rm BH} \propto M_{\rm *,sph}^{2-3}$ and $M_{\rm BH}/M_{\rm *,sph}$ equal to $0.02\%$ at $M_{\rm BH} \approx 10^6~\rm M_\odot$, 
indicating that gas-rich processes feed the black hole more efficiently than the host bulge as they coevolve. 
We additionally report that: i) our S\'ersic galaxy sample follows $M_{\rm BH} \propto M_{\rm *,sph}^{1.48 \pm 0.20}$, 
a less steep sequence than previously reported; 
ii) bulges with S\'ersic index $n_{\rm sph}<2$, argued by some to be pseudo-bulges, 
are not offset to lower $M_{\rm BH}$ from the correlation defined by the current bulge sample with $n_{\rm sph}>2$; 
and iii) $L_{\rm sph}$ and $L_{\rm gal}$ correlate equally well with $M_{\rm BH}$, in terms of intrinsic scatter, only for early-type galaxies 
-- once reasonable numbers of spiral galaxies are included, the correlation with $L_{\rm sph}$ is better than that with $L_{\rm gal}$.

@@

Here we investigate in more detail 

@@

\subsection{Black hole mass -- spheroid stellar mass}
Finally, we present the $M_{\rm BH} - M_{\rm *,sph}$ diagram in Figure \ref{fig:mbhmasssph}, 
and its linear regression analysis in Table \ref{tab:lregmass}. 
The bulges of the early-type galaxies follow $M_{\rm BH} \propto M_{\rm *,sph}^{1.04 \pm 0.10}$,
consistent with a dry-merging formation scenario\footnote{In dry mergers, the black hole and the bulge grow at the same pace, 
increasing their mass in lock step.},
and define a tight \emph{early-type sequence} with intrinsic scatter $\epsilon_{(Y|X)} = 0.43 \pm 0.06\rm~dex$. 
\cite{graham2012bent} reported that the $M_{\rm BH}/M_{\rm dyn,sph}$ ratio for core-S\'ersic galaxies was 0.36\% 
($M_{\rm dyn,sph}$ is the spheroid dynamical mass), 
and discussed the many implications of this.  
Using a larger data sample, \cite{grahamscott2013} reported that the $M_{\rm BH}/M_{\rm *,sph}$ ratio was 0.49\% for core-S\'ersic galaxies.   
Here we find a median $M_{\rm BH}/M_{\rm *,sph}$ ratio of $0.50 \pm 0.04\%$ for the 22 core-S\'ersic galaxies 
and $0.68 \pm 0.04\%$ for the 45 early-type galaxies.  
Among other things, this higher value (previously reported to be $0.1 - 0.2\%$ for all galaxy types, e.g.~\citealt{marconihunt2003}), 
boosts estimates of the black hole mass function and mass density based on galaxy/spheroid luminosity functions. \\
The bulges of the spiral galaxies trace a steeper \emph{late-type sequence}, 
whose slope is less well constrained due to the smaller size of the subsample and, more importantly, 
to the smaller range in $M_{\rm *,sph}$ that the subsample spans. 
For the bulges of spiral galaxies, the BCES code returns a log-linear relation with a slope $= 3.00 \pm 1.30$, 
while the modified FITEXY routine finds a shallower (but still consistent within the $1\sigma$ uncertainty) 
slope $= 2.28^{+1.67}_{-1.01}$.
More data would be welcome to better constrain the slope of this \emph{late-type sequence}, 
although we note that direct measurements of black hole masses below $10^6\rm~M_\odot$ are extremely challenging to obtain 
with the current technological resources. 
In this regard, using a sample of $\sim$140 low-redshift ($z \leq 0.35$, with a median redshift $\langle z \rangle = 0.085$) 
bulges hosting Active Galactic Nuclei (AGNs) with virial
black hole masses $10^5 \lesssim M_{\rm BH}/{\rm M_\odot} \lesssim 2 \times 10^6$ \citep{jiang2011a}, 
\cite{grahamscott2015} showed that they roughly follow the quadratic $M_{\rm BH} - M_{\rm *,sph}$ relation defined by their S\'ersic bulges.
The majority of our spiral galaxies host an AGN\footnote{According to the nuclear classification reported on NED 
(NASA Extragalactic Database), among our 17 spiral galaxies, at least 12 host a Seyfert AGN and one hosts a LINER AGN.} and
we anticipate here that the correlation traced by our spiral galaxy bulges 
may track the location of these lower mass AGN in the $M_{\rm BH} - M_{\rm *,sph}$ diagram.
That is, the AGNs appear to be the low-mass continuation of our tentative \emph{late-type sequence} shown in Figure \ref{fig:mbhmasssph} 
and this will be explored with more rigour in a forthcoming paper. \\

{\bf 
As a final remark, we comment on the work by \cite{reinesvolonteri2015}, 
who investigated the relationship between black hole mass and total galaxy stellar mass, $M_{\rm *,gal}$. 
Their Figure 8 presents the $M_{\rm BH} - M_{\rm *,gal}$ distribution 
for a sample of $\approx 260$ local AGNs with virial black hole masses  
and for $\approx 80$ galaxies with dynamical black hole masses. 
They concluded that the AGN sample and the early-type galaxies with quiescent black holes define two distinct sequences 
in their $M_{\rm BH} - M_{\rm *,gal}$ diagram; 
these two sequences have similar slope, but a normalization factor different by more than one order of magnitude.  
Since we noted that the \cite{jiang2011a} AGN sample follows the steeper $M_{\rm BH} - M_{\rm *,sph}$ correlation 
traced by our spiral galaxy bulges (the majority of which host an AGN), 
it would be interesting to recover the $M_{\rm BH} - M_{\rm *,sph}$ distribution 
also for the AGN sample of \cite{reinesvolonteri2015}. \\
However we do note that there is emerging evidence (e.g.~\citealt{busch2015,subramanian2015}) 
for a population of bulges with black hole masses residing below (or to the right of) the red and blue $M_{\rm BH} - M_{\rm *,sph}$ 
sequences constructed here using samples with directly measured black hole masses, as speculated by Batcheldor (2010).
}

@@

\begin{itemize}
\item The logarithmic $M_{\rm BH} - M_{\rm *,sph}$ relation for the spheroidal components of early-type (elliptical + lenticular) galaxies 
      has a slope of $1.04 \pm 0.10$ and intrinsic scatter $\epsilon_{(Y|X)} = 0.43 \pm 0.06\rm~dex$. 
      We call this tight correlation an \emph{early-type sequence}. 
      The $M_{\rm BH} - M_{\rm *,sph}$ log-relation for the bulges of late-type (spiral) galaxies has a slope of $2-3$,
      which is less well constrained due to
      the smaller size of the subsample and, more importantly, the smaller range in spheroid stellar mass 
      ($3 \times 10^9 \lesssim M_{\rm *,sph}/{\rm M_\odot} \lesssim 3 \times 10^{10}$) that the subsample spans. 
      We refer to this correlation as a \emph{late-type sequence}. 
      In (gas-poor) early-type galaxies, the black hole and the stellar content of the spheroidal component grow at the same pace, 
      following a linear $M_{\rm BH} - M_{\rm *,sph}$ relation. 
      In (gas-rich) spiral galaxies, the black hole grows faster than its host bulge, 
      following a quadratic/cubic $M_{\rm BH} - M_{\rm *,sph}$ relation. 
      Unsurprisingly, in a color-magnitude diagram\footnote{Total $B-V$ colors, 
      corrected for inclination, Galactic extinction and K-correction, 
      were taken from the HyperLEDA online database \citep{hyperleda}.}, 
      our early- and late-type galaxies occupy the two distinct regions of the red sequence and the blue cloud, respectively. 
      For analogy with this, we refer to our \emph{early-type sequence} as a \emph{red sequence} 
      and to our \emph{late-type sequence} as a \emph{blue sequence}. 
\item The median $M_{\rm BH}/M_{\rm *,sph}$ ratio for the early-type galaxies is $0.68 \pm 0.04\%$. 
      This value is dramatically larger than what was previously reported ($0.1 - 0.2\%$ for all galaxy types, e.g.~\citealt{marconihunt2003}), 
      but in close agreement with the value of $0.49\%$ reported by \cite{grahamscott2013} for core-S\'ersic spheroids. 
\item The logarithmic $M_{\rm BH} - M_{\rm *,sph}$ relations for the core-S\'ersic and S\'ersic spheroids 
      have slopes with over-lapping uncertainties ($1.19 \pm 0.23$ and $1.48 \pm 0.20$, respectively).  
      The S\'ersic relation is less steep than, but also has over-lapping uncertainties with, 
      the slope of $2.22 \pm 0.58$ reported by \cite{scott2013} for S\'ersic spheroids.  
      The distinction between core-S\'ersic and S\'ersic spheroids found by \cite{scott2013} is thus less pronounced here. 
\item In the $M_{\rm BH} - L_{\rm sph}$ (or $M_{\rm BH} - M_{\rm *,sph}$) diagram, for early-type galaxies, 
      we did not observe the change in slope required for consistency with the log-linear $M_{\rm BH} - \sigma$ and bent $L_{\rm sph} - \sigma$ correlations.
      This issue of inconsistency remains therefore an open question. 
      It might be that we are still not probing enough low-mass black holes ($M_{\rm BH} < 10^7\rm~M_\odot$) 
      for the subsample of early-type galaxies, 
      or that the transition from  $L_{\rm sph} \propto \sigma^2$ at low luminosities to $L_{\rm sph} \propto \sigma^{(5-6)}$ at high luminosities 
      is less sharp than previously thought.
      We intend to investigate this point in our future work. 
\item {\bf It has been argued that pseuso-bulges (disk-like, rotation-supported systems, built from secular processes) 
      do not follow the $M_{\rm BH} - L_{\rm sph}$ correlation defined by classical bulges 
      (spheroidal, pressure-supported systems, formed through violent processes).  
      The recent literature (e.g.~\citealt{sani2011,beifiori2012}) has distinguished between pseudo- and classical bulges 
      according to their S\'ersic index, $n_{\rm sph}$. 
      Although we do not consider the S\'ersic index a good indicator of the nature of a bulge (e.g.~\citealt{grahamworley2008}), 
      we investigated this point and found that, within the current dataset,  
      spheroids with $n_{\rm sph}<2$ 
      are not offset to lower $M_{\rm BH}$ from the $M_{\rm BH} - L_{\rm sph}$ correlation defined by spheroids with $n_{\rm sph}>2$. }
\item The $M_{\rm BH} - L_{\rm gal}$ and $M_{\rm BH} - L_{\rm sph}$ correlations have the same level of intrinsic scatter 
      when considering early-type galaxies only. 
      Once reasonable numbers of spiral galaxies are included, 
      $M_{\rm BH}$ correlates better with $L_{\rm sph}$ than with $L_{\rm gal}$ (see also \citealt{erwingadotti2012,beifiori2012}). 
\end{itemize}

@@

\acknowledgments
GS warmly thanks Luca Cortese, Elisabete Lima Da Cunha, Duncan Forbes and Gonzalo Diaz for useful discussion. 
We also thank the anonymous referee for useful comments and suggestions. 
This research was supported by Australian Research Council funding through grants
DP110103509 and FT110100263.
This work is based on observations made with the IRAC instrument \citep{fazio2004IRAC} 
on-board the Spitzer Space Telescope, 
which is operated by the Jet Propulsion Laboratory, 
California Institute of Technology under a contract with NASA.
This research has made use of the GOLDMine database \citep{goldmine} and the NASA/IPAC Extragalactic Database (NED) 
which is operated by the Jet Propulsion Laboratory, California Institute of Technology, 
under contract with the National Aeronautics and Space Administration. 
We acknowledge the usage of the HyperLeda database (\url{http://leda.univ-lyon1.fr}).
The BCES routine \citep{akritasbershady1996} was run via the python module 
written by Rodrigo Nemmen \citep{nemmen2012}, which is available at \url{https://github.com/rsnemmen/BCES}.

### 2015-11-20 papers/MbhMsph/arxiv/mm.tex
victory2

{\bf We therefore do not assume that all bulges with $n_{\rm sph}<2$ are built from internal processes in the disk 
(i.e.~are what some authors call pseudo-bulges). } \\

@@

{\bf 
As a final remark, we comment on the work by \cite{reinesvolonteri2015}, 
who investigated the relationship between black hole mass and total galaxy stellar mass, $M_{\rm *,gal}$. 
Their Figure 8 presents the $M_{\rm BH} - M_{\rm *,gal}$ distribution 
for a sample of $\approx 260$ local AGNs with virial black hole masses  
and for $\approx 80$ galaxies with dynamical black hole masses. 
They concluded that the AGN sample and the early-type galaxies with quiescent black holes define two distinct sequences 
in their $M_{\rm BH} - M_{\rm *,gal}$ diagram; 
these two sequences have similar slope, but a normalization factor different by more than one order of magnitude.  
Since we noted that the \cite{jiang2011a} AGN sample follows the steeper $M_{\rm BH} - M_{\rm *,sph}$ correlation 
traced by our spiral galaxy bulges (the majority of which host an AGN), 
it would be interesting to recover the $M_{\rm BH} - M_{\rm *,sph}$ distribution 
also for the AGN sample of \cite{reinesvolonteri2015}. \\
However we do note that there is emerging evidence (e.g.~\citealt{busch2015,subramanian2015}) 
for a population of bulges with black hole masses residing below (or to the right of) the red and blue $M_{\rm BH} - M_{\rm *,sph}$ 
sequences constructed here using samples with directly measured black hole masses, as speculated by Batcheldor (2010).
}

@@

\item {\bf It has been argued that pseuso-bulges (disk-like, rotation-supported systems, built from secular processes) 

@@

      are not offset to lower $M_{\rm BH}$ from the $M_{\rm BH} - L_{\rm sph}$ correlation defined by spheroids with $n_{\rm sph}>2$. }

>>>>

We therefore do not assume that all bulges with $n_{\rm sph}<2$ are built from internal processes in the disk 
(i.e.~are what some authors call pseudo-bulges).  \\

@@

As a final remark, we comment on the work by \cite{reinesvolonteri2015}, 
who investigated the relationship between black hole mass and total galaxy stellar mass, $M_{\rm *,gal}$. 
Their Figure 8 presents the $M_{\rm BH} - M_{\rm *,gal}$ distribution 
for a sample of $\approx 260$ local AGNs with virial black hole masses  
and for $\approx 80$ galaxies with dynamical black hole masses. 
They concluded that the AGN sample and the early-type galaxies with quiescent black holes define two distinct sequences 
in their $M_{\rm BH} - M_{\rm *,gal}$ diagram; 
these two sequences have similar slope, but a normalization factor different by more than one order of magnitude.  
Since we noted that the \cite{jiang2011a} AGN sample follows the steeper $M_{\rm BH} - M_{\rm *,sph}$ correlation 
traced by our spiral galaxy bulges (the majority of which host an AGN), 
it would be interesting to recover the $M_{\rm BH} - M_{\rm *,sph}$ distribution 
also for the AGN sample of \cite{reinesvolonteri2015}. 
However we do note that there is emerging evidence (e.g.~\citealt{busch2015,subramanian2015}) 
for a population of bulges with black hole masses residing below (or to the right of) the red and blue $M_{\rm BH} - M_{\rm *,sph}$ 
sequences constructed here using samples with directly measured black hole masses, as speculated by Batcheldor (2010).

@@

\item It has been argued that pseuso-bulges (disk-like, rotation-supported systems, built from secular processes) 

@@

      are not offset to lower $M_{\rm BH}$ from the $M_{\rm BH} - L_{\rm sph}$ correlation defined by spheroids with $n_{\rm sph}>2$. 
